---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# vectorTools


The goal of vectorTools is to provide a common set of functions used to analyze data on vector, a local compute server at the institute for medical epidemiology and biostatistics.

## Installation

You can install the development version of vectorTools like so:

``` r
library(devtools)
install_github("ararder/vectorTools")
```

## UK Biobank
On vector, phenotype for UKB is stored in  separate files. The files are
very large, so it is not feasible to read the entire dataset in, and then drop the variables you dont
want. 

For this, there is the ukb_extract() function, which takes as argument a vector of fields to extract.
The function will first see if the field exists in each of the datasets, and then extract it if it exists.


```{r, eval = FALSE}
library(tidyverse)
library(vectorTools)
library(data.table)

# Extracts field f.34
## Note the "\\. - this is to ensure an exact match
tt <- ukb_extract("f.34\\.")

selfrep_data <- ukb_extract("f.20002.")
cols <- fread("/nfs/AGE/UKB/data/180524/r/ukb22140.tab", nrows = 0)
```


The matching of each character vector to the fields is with stringr::str_detect
This means that you have to be careful if the field id can match several ideas, for example:
```{r}
field_to_get <- c("f.34")
made_up_fields <- c("f.34", "f.34100", "f.314")

# This also matches f.34100 !!
stringr::str_detect(made_up_fields, "f.34")

# To handle this, add in the \\. to also match the ., which correponds to end of field id-
stringr::str_detect(made_up_fields, "f.34\\.")

```
In general, UKB field ids can be mapped to the showcase ids as following:
field [20446](https://biobank.ndph.ox.ac.uk/showcase/field.cgi?id=20446) becomes
f.20446.[subfield1].[subfield2]

How many subfields there are will depend on each field. Generally the contain cases where
the phenotype is measured several times per visit, and if the phenotype has been measured at more than one visit.

# Extracting ICD definitions from UKB
A common task is to get the list of individuals with atleast one instance of an ICD code.
If you only care about whether an individual has ever been diagnosed with a disorder, the field to use is
[41270](https://biobank.ndph.ox.ac.uk/showcase/field.cgi?id=41270)

In UKB, many fields are stored in a wide format - you have many columns for each individual, and the code could by in any of these columns. For this i use check_for_code, which checks if the specified code is is any of the columns in the dataset provided (except column 1, which is assumed to be id)
```{r, eval = FALSE}
library(tidyverse)
library(vectorTools)

# assuming you have read permission for UKB
# list of all unique icd codes

# first we extract f.41270, then extract all individuals with F43
icd_data <- ukb_extract("f.41270")
anxiety <- check_for_code(codes = "F43", col_name = "anxiety", data = icd_data)

# What if we want anxiety AND mdd? - wrap multiple icd with c(code1, code2)
mdd_anx <- check_for_code(codes = c("F32","F43"), col_name = "anx_mdd", data = icd_data)

# check_for_code can easily be extended to several groups, and if you want to use
# multiple ICD codes to define a single disorder
codes <- list(
  bip =  c("F30", "F31"),
  mdd_recur = c("F33"),
  scz = c("F20", "F25"),
  mdd_single = c("F32")
)

cases <- map2_df(codes, names(codes), check_for_code, data = icd_codes)
```

# Extracting selfreported definitions from UKB
Since datafield 20002 - (selfreported diagnosis) follows the same long format 
as 41270, we can easily extend the above code for selfreported codes

By default, ukb_extract removes individuals whom have withdrawn their consnet
```{r, eval=FALSE}
selfrep_data <- ukb_extract("f.41270")

# 1287 corresponds to the selfrep code for anxiety in field 20002
anxiety <- check_for_code(codes = "1287", col_name = "anxiety", data = selfrep_data)

codes <- list(
  anxiety_panic_attacks = c("1287"),
  nervous_breakdown = c("1288"),
  mig_or_epilipsy = c("1290",	"1265")
)

cases <- map2_df(codes, names(codes), check_for_code, data = selfrep_data)
```

# Crime 3

```{r example, eval =FALSE}
library(vectorTools)
# Establish connection to Crime3
con <- crime3_con()

# set the ICD codes you are interested in
diagnosis_codes <- c("F25", "F29")

# extract all observations any of those codes from specialist care
scz_dia <- crime3_icd(diagnosis_codes)

# or from primary care
scz_dia_primary <- crime3_icd(diagnosis_codes, specialist=FALSE)

# For each individual, find the first diagnosis of each unique diagnosis
# with 3 digits codes like F3222A becomes F32. 
scz_first_dia <- crime3_first_unique_diagnosis(scz_dia, icd_digits = 3)

# we can increase the number of digits for each unique diagnosis, so that we find
# the first occurence of both F323 and F322 if the individual has both.
scz_first_dia_sub <- crime3_first_unique_diagnosis(scz_dia, icd_digits = 4)

## basic example code
```


